---
title: "Chad Huntebrinker's Assignment 3"
author: "Chad Huntebrinker"
date: "2025-09-07"
output:
  word_document: default
  html_document: default
---

```{r setup}
library(ggplot2)
library(tidyr)
```

Exercise 1:
```{r}
#Exercise 1
#Explore the dataset
head(mpg)
str(mpg)
dim(mpg)

#Convert the character
character_columns <- names(mpg)[sapply(mpg, is.character)]
mpg[character_columns] <- lapply(mpg[character_columns], as.factor)
str(mpg)

#Summarize
summary(mpg)

#We're converting cyl to a factor (since the number of cyl is either 4, 6, 8)
mpg$cyl <- lapply(mpg$cyl, as.factor)
```

Exercise 2:
See if there are any missing values for cty and find the most frequent value for cty. Also find the mean and median values for cty too.
```{r}
#Exercise 2
#Missing values for cty
sum(is.na(mpg$cty))

#Finding the most frequent cty value
table(mpg$cty)

mean(mpg$cty)
median(mpg$cty)
```

There are no missing values for cty, 18 is the most frequent value, and the mean and median are close to the value of 18.

Find the peak values of highway MPG and get the mean and median:
```{r}
#Find the peak values of hwy
table(mpg$hwy)


#mean and median of hwy
mean(mpg$hwy)
median(mpg$hwy)
```

Looks like the peak values are 17 and 26. The reason why there are multiple values for hwy is because cars with a lower fuel efficiency (larger cars like SUVs and trucks would be in this category) are around 17mpg and cars with a higher fuel efficiency (like sedans and smaller cars) are around 26mpg.

The mean and median does show some useful information, like what the average mpg of all cars are. But it doesn't show the full story (like how there are two peaks) like the table and the histogram will show.

Create the two histograms of city MPG and highway MPG and make the subplots of one plot:
```{r}
#Create histograms
p1 <- ggplot(mpg, aes(x=cty)) +
  geom_histogram(fill="red") +
  labs(
    title = "Count of City MPG",
    x = "MPG in City",
    y = "Count"
  ) +
  theme_minimal()

p2 <- ggplot(mpg, aes(x=hwy)) +
  geom_histogram(fill="blue") +
  labs(
    title = "Count of Highway MPG",
    x = "MPG on Highway",
    y = "Count"
  ) +
  theme_minimal()

first_row <- cowplot::plot_grid(p1, nrow = 1)
second_row <- cowplot::plot_grid(p2, nrow = 1)

cowplot::plot_grid(p1, p2, ncol = 1, nrow = 2)
```


Get an area plot after transforming the data:
```{r}
#Transform data from wide to long and then creat an area plot of MPG in city and highway
mpg_long <- mpg %>% 
  pivot_longer(cols = c(cty, hwy),
               names_to = "mileage_type",
               values_to = "mpg")


ggplot(mpg_long, aes(x = mpg, fill = mileage_type)) +
  geom_area(stat = "bin", binwidth = 1, alpha = 0.5, position = "identity") +
  labs(
    title = "City and Highway MPG",
    x = "MPG",
    y = "Count",
    fill = "Mileage Type"
  ) +
  theme_minimal()
```

